\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-04-03 eg6-letters/NIM/EG6_RTPC_NIM.tex
Correction section 1 of NIM

\begin{abstract}
A new Radial Time Projection Chamber (RTPC) was developed at Jefferson 
Laboratory to track low-energy nuclear recoils for the purpose of measuring
exclusive nuclear channels, such as coherent Deeply Virtual Compton Scattering
and coherent meson production on $^4$He. In such processes, the $^4$He nucleus
remains intact and recoils as a whole in the final state. The CEBAF Large
Acceptance Spectrometer (CLAS) has been used for similar studies on proton
targets, however it cannot track recoils with momenta less than 250 MeV/c. In
2009, we carried out measurements using the CLAS spectrometer supplemented by
the RTPC positioned directly around a gaseous $^4$He target, allowing a detection
threshold as low as $~$100  MeV/c for $^4$He. This article discuses the design,
work principle, calibration methods and performances of the RTPC.
\end{abstract}

@@

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a hadron) necessitates some adaptations
in this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles where the acceptance of the EC is poor. To extend the range,
an additional Inner Calorimeter (IC) was built for the E01-113 experiment in 2005.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long, corresponding
to 17 radiation lengths. The achieved energy resolution is around 3 to 4$\%$ for photon
energies between 2 GeV and 5 GeV, with an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}.

%M{\o}ller electrons can be
%very problematic at such small angles, so a 5 Tesla solenoid magnet has been
%added around the target as a magnetic shield against this low energy background.
For coherent nuclear processes, the RTPC was developed to extend the tracking
range of CLAS to low energy nuclei and was used during a one month experimental run
in 2009 with a longitudinally polarized electron beam of energy 6.064 GeV incident 
on a gaseous $^{4}$He target.
%, with the goal of measuring asymmetries in Deep Virtual Compton 
%Scattering (DVCS)~\cite{proposal} and exclusive meson production.

Moreover, a 5~T solenoid was placed around the RTPC-target system to shield the IC
and RTPC from the large flux of low energy M{\o}ller electrons. The solenoid field
serves also as the analyzing magnet for charged particle tracking by the RTPC.

In this paper we report the design, calibration, and performance of the radial TPC,
organized as follows.  Section \ref{prev_tpc} presents an overview of the history of
TPCs.  Section \ref{sec_design} details the physics considerations for our RTPC's
design and internal structure. Section \ref{sec_readout} describes the properties
of the read-out system.  Calibration strategies are discussed in section \ref{sec_calib}.
Finally, the performance of the RTPC is described in section \ref{sec_perf}.

\section{TPC technology} \label{prev_tpc}
In the late 1970s, David R. Nygren presented the concept of the Time Projection 
Chambers (TPCs) as a new tool to track charged particles based on combining 
electric and magnetic fields within a sensitive gas volume. The ionized 
electrons are collected by a position-sensitive electron collection system.  
Later in 1997, F.~Sauli introduced the Gas Electron Multiplier (GEM) as an 
effective gas-gain element to improve the tracking efficiency of the TPCs 
\ref{gem_sauli}. Presently, TPCs are widely used in physics experiments, 
such as high-energy physics (ALEPH, DELPHI), heavy-ion physics (STAR, ALICE) 
and rare-event searches (XENON, NEXT).

The first implementation of a radial TPC was performed by the BoNuS experiment
\cite{BONUS} in 2005. The success of the BoNuS-RTPC in detecting the slow protons
motivated the development of this new RTPC. For the channels of interest in
CLAS-EG6 experiment, the recoiling $^{4}$He nuclei have momentum per charge
around 150 MeV/c. The detection of these low-energy nuclei was the primary
motivation for designing this detector.

>>>>

\begin{abstract}
A new Radial Time Projection Chamber (RTPC) was developed at Jefferson 
Laboratory to track low-energy nuclear recoils for the purpose of measuring
exclusive nuclear channels, such as coherent Deeply Virtual Compton Scattering
and coherent meson production on $^4$He. In such processes, the $^4$He nucleus
remains intact and recoils as a whole in the final state. The CEBAF Large
Acceptance Spectrometer (CLAS) has been used for similar studies on proton
targets, however it cannot track low energy $\alpha$ particles. In
2009, we carried out measurements using the CLAS spectrometer supplemented by
the RTPC positioned directly around a gaseous $^4$He target, allowing a detection
threshold as low as $~$150  MeV/c for $^4$He. This article discuses the design,
work principle, calibration methods and performances of the RTPC.
\end{abstract}

@@

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a hadron) necessitates an upgrade
of this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles, where the acceptance of the EC is poor. To extend the detection range,
an inner calorimeter (IC) was built for the E01-113 experiment in 2005.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long, corresponding
to 17 radiation lengths. The achieved energy resolution is around 3 to 4$\%$ for photon
energies between 2 GeV and 5 GeV, with an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}. To protect this detector from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid was placed around the target to shield the IC. 
In addition, the recoil nuclei of the coherent DVCS on helium was detected 
using a RTPC developed to track low energy nuclear fragments. The solenoid field
serving as the analyzing magnet for the particle tracking in the RTPC. This 
setup was used during a three months experimental run
in 2009 with a longitudinally polarized electron beam of 6.064 GeV incident 
on a gaseous $^{4}$He target.

In this paper we report the design, calibration, and performance of the radial TPC,
organized as follows. Section \ref{sec_design} details the physics considerations for our RTPC's
design and internal structure. Section \ref{sec_readout} describes the properties
of the read-out system. Calibration strategies are discussed in section \ref{sec_calib}.
Finally, the performance of the RTPC is described in section \ref{sec_perf}.
